\section{Vector Analysis}
\begin{itemize}
    \item \textbf{Triple scalar product}:
          \begin{equation*}
              \mathbf{A}\cdot(\mathbf{B} \times \mathbf{C}) =
              \left|\begin{matrix}
                  A_x & A_y & A_z \\
                  B_x & B_y & B_z \\
                  C_x & C_y & C_z
              \end{matrix}\right|
          \end{equation*}
          Note:
          \begin{itemize}
              \item Performing an even number of swaps on the determinant results in the same answer, whereas performing an odd number of swaps simply results in an answer of the opposite sign.
              \item Geometrically, the triple scalar product represents the volume of a parallelopiped formed by $\mathbf{A}$, $\mathbf{B}$, and $\mathbf{C}$.
          \end{itemize}
    \item \textbf{Triple vector product}:
          \begin{equation*}
              \mathbf{A} \times (\mathbf{B} \times \mathbf{C})=(\mathbf{A} \cdot \mathbf{C})\mathbf{B} - (\mathbf{A} \cdot \mathbf{B})\mathbf{C}
          \end{equation*}
          To remember this formula, the right hand side is a linear combination of the vectors in the parentheses on the left hand side. The coefficients are the outside vector dotted with each of the inside vectors. Both terms have $\mathbf{A}$, $\mathbf{B}$, and $\mathbf{C}$ exactly once. Finally, the positive term on the right hand side belongs to the vector in the middle on the left hand side ($\mathbf{B}$ in this case).
    \item Vector properties
    \item Torque??
    \item Line integrals pg. 320, conservative vector fields pg. 322 and curl, gradient
    \item pg. 323 conservative -> potential
    \item pg. 325 vecAdvecA = AdA
    \item pg. 326 exact diff (and mixed partial equal) when conservative
    \item pg. 330 Green's Theorem
    \item pg. 332 tanget and outward normal vec
    \item pg. 332 9.13 9.14 divergence theorem
    \item pg. 333 9.18 9.19 Stokes' Theorem
\end{itemize}